TODO: A Survey of StarCraft AI Techniques

### 2012-05-11 survey.tex
first pass on state-of-the-art

{\color{blue}
perhaps TODO ``how it's done in the industry'' regrouping the first paragraph of each of the following 3 sections?
}
%The most common decision making are Decision Trees. Some of the advantages of the decision trees are: simple to understand and interpret acting like a white box where we can explain why a decision is taken, and they are compatible with other decision techniques. For complex decisions we have many algorithms to generate an optimum decision tree like ID3 or C4.5, depends on the problem we will use one or another.

\subsection{Strategy}
%XXX Every AI needs to observe the world in order to analyze it and take decisions. In RTS games this task isn't trivial since we have to deal with imperfect information.

@@

In research, the main methods used for strategic decision-making are case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}. Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). Onta\~{n}\'{o}n et al. \cite{CBR_Planning} base their real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR learned from expert demonstrations on Wargus. They update their previous CBP approach by using situation assessment for retrieving the ``right'' plan, and improve the speed of CBP by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). Hoang et al. \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN. %Madeira et al. \cite{Madeira06}

@@

Most commercial games avoid gathering this information by using perfect information all the time (cheating). Sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. \{\color{blue}XXX (perhaps move it above, to the start of subsection}

@@

\subsection{Tactics}
Terrain analysis supplies the AI with chunks of abstract information about the map in order to help in making decisions. This analysis is usually performed off-line, in order to save processing time during gameplay. Pottinger \cite{Pottinger00} described the \emph{BANG} engine implemented by Ensemble Studios for the game Age of Empires II. This engine provides terrain analysis functionalities to the game using influence maps and areas with connectivity information. Forbus et al. \cite{Forbus2002} showed the importance to have qualitative spatial information for wargames, for which they used geometric and pathfinding analysis. Hale et al. \cite{Hale08} presented a 2D geometric navigation mesh generation method from expanding convex regions from seeds. Finally, Perkins \cite{Perkins10} applied Voronoi decomposition (then pruning) %\cite{Karavelas04}
to detect regions and relevant choke points in RTS maps. This approach is implemented for StarCraft in BWTA\footnote{\url{http://code.google.com/p/bwta/}}.
% Isla 2006 Game AI Programming Wisdom 3. Charles River Media. chapter Probabilistic Target Tracking and Search Using Occupancy Maps, 379–388. ?

Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in FPS units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. 
% comparable with human experts ?
Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in Tankbattle. XXX 
Weber et al. \cite{WeberAIIDE11} used a particle model with a simple linear trajectory update to track opponent units under fog of war in StarCraft.

\subsection{Units Control}

% {\color{ForestGreen}
% Explain centralized models (like blackboard) and distributed/hierarchically models.
% }
The most common pathfinding algorithm is A*, but the big problem of this algorithm is the time and  memory consumption, a show-stopper in a real-time environment.
To solve this the search space needs to be reduced as much as possible. One common technique is reducing the tile map search into a navigation graph map. To do this some polygonal triangulations are calculated and abstracted on a grid-based map and this is the TRA* (Triangulation Reduction A*) algorithm\cite{Demyen_2006}.

Craig Reynolds introduced algorithmic steering behaviors to simulate the aggregate motion of swarms using procedural animation\cite{Reynolds_1999}. His model describes how a system of multiple agents, each steering according to simple local rules, can produce the collective movement found in flocks of birds. With this behaviors, we are able to simulate intelligent squad movements and do tactical moves to flank the opponent in a RTS game\cite{Danielsiek_2008,SynnaeveMicroCig11}, improving the survival of our units.

A good alternative to Flocking is using Reciprocal Velocity Obstacles (RVO)\cite{Berg_2008}, since it is easier to control the behavior and they guarantee to prevent collisions. With this algorithm, we can move a large squad crossing another squad and avoid all possible collisions between units. This type of techniques are already applied in commercial RTS games like Starcraft 2 or Supreme Commander 2 with great success.

Another popular technique is Potential Fields. basically, Potential Fields work like a charged particle moving through a magnetic field. The process of tuning and optimizing the parameters of potential field charges can be time reduced applying learning algorithms such as Q-Learning\cite{Liu_2008}. Another common issue we have to deal with when using potential fields is the local optima problem that can stuck the unit away from the goal. Some examples of the utilities of potential files in RTS games are avoiding obstacles (navigation), avoiding opponent fire, or staying at maximum shooting distance\cite{Hagelback08,Hagelback09}.

There are some interesting uses of reinforcement learning (RL) \cite{Sutton}: \cite{Marthi05} employs concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. \cite{Madeira06} advocates the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. In real game setups, the actions spaces to explore is gigantic, it requires to use the structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. There are other was to learn parameters of an underlying model, for instance \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. The difficulty to work with multi-scale goals and plans is handled directly by case-based reasoning (CBR), which has been adapted for units behavior with continuous action models \cite{Molineaux08}, an integrated RL/CBR algorithm using continuous models, or with hybrid CBR/RL transfer learning \cite{CBR-RL}. Reactive planning (for behavior languages) \cite{WeberCig10}, a decompositional planning similar to hierarchical task networks \cite{HTNPlanning}, allows for plans to be changed at different granularity levels and so for multi-scale (hierarchical) goals integration.

Self-organizing-maps (SOM) learned with evolutionary algorithms provide some way to learn (realistic) parameters for RTS units behavior \cite{teamCompositionRTS}. Bayesian programming has been applied to inverse fusion of the sensory inputs of the units \cite{SynnaeveMicroCig11}, which subsumes potential fields, allowing for integration of tactical goals directly in micro-management. There are also some cognitive approaches to RTS AI \cite{SORTS}, in which there is a human-like attention model for micro-management.

>>>>

%{\color{blue}
%perhaps TODO ``how it's done in the industry'' regrouping the first paragraph of each of the following 3 sections?
%}

\subsection{Strategy}
%Every AI needs to observe the world in order to analyze it and take decisions. In RTS games this task isn't trivial since we have to deal with imperfect information.
%The most common decision making are Decision Trees. Some of the advantages of the decision trees are: simple to understand and interpret acting like a white box where we can explain why a decision is taken, and they are compatible with other decision techniques. For complex decisions we have many algorithms to generate an optimum decision tree like ID3 or C4.5, depends on the problem we will use one or another.

@@

In research, the main methods used for strategic decision-making are case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}. Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). Onta\~{n}\'{o}n et al. \cite{CBR_Planning} base their real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR learned from expert demonstrations on Wargus. They update their previous CBP approach by using situation assessment for retrieving the ``right'' plan, and improve the speed of CBP by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). Hoang et al. \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. %Madeira et al. \cite{Madeira06}

@@

Most commercial games avoid gathering this information by using perfect information all the time (cheating). Sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. {\color{blue}XXX (perhaps move it above, to the start of subsection}

@@

\subsection{Tactics}
Units have different abilities, which leads to different possible tactics. Most often, fast or invisible or flying units are not cost-effective in head-to-head fights against slower bulky units. Taking good tactical decisions is taking advantage of the terrain and positions of the different units (in a timely manner). Terrain analysis supplies the AI with structured information about the map in order to help making decisions. This analysis is usually performed off-line, in order to save CPU time during the game. Pottinger \cite{Pottinger00} described the \emph{BANG} engine implemented by Ensemble Studios for the game Age of Empires II. This engine provides terrain analysis functionalities to the game using influence maps and areas with connectivity information. Forbus et al. \cite{Forbus2002} showed the importance to have qualitative spatial information for wargames, for which they used geometric and pathfinding analysis. Hale et al. \cite{Hale08} presented a 2D geometric navigation mesh generation method from expanding convex regions from seeds. Finally, Perkins \cite{Perkins10} applied Voronoi decomposition (then pruning) %\cite{Karavelas04}
to detect regions and relevant choke points in RTS maps. This approach is implemented for StarCraft in BWTA\footnote{\url{http://code.google.com/p/bwta/}}.
% Isla 2006 Game AI Programming Wisdom 3. Charles River Media. chapter Probabilistic Target Tracking and Search Using Occupancy Maps, 379–388. ?

Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in Tankbattle. Weber et al. \cite{WeberAIIDE11} used a particle model with a simple linear trajectory update to track opponent units under fog of war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{Chung05} adapted Monte-Carlo tree search (MCTS) to planning in RTS games and applied it to a capture-the-flag mod of Open RTS. Balla and Fern \cite{UCT} applied upper confidence bounds on trees (UCT: a MCTS algorithm) to tactical assault planning in Wargus. On influence maps, \cite{teamCompositionRTS} studied team composition and maneuvering by learning a self-organizing map, while \cite{HagelbackJ08} presented a multi-agent potential field based bot. Finally, Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial (tactical) reasoning in RTS games.

\subsection{Units Control}
Units control, also called micro-management in RTS games, aims at maximizing the effectiveness of units, it starts with pathfinding and goes up to simultaneous control of units of different types in complex battles on heterogeneous terrain. The most common pathfinding algorithm is A*, but its big problem is CPU time and memory consumption, hard to satisfy in a complex, dynamic, real-time environment on multiple units. %To solve this the search space needs to be reduced as much as possible. 
A common technique consists in reducing the search into a navigation graph map. Among others, Triangulation Reduction A* computes polygonal triangulations on a grid-based map \cite{Demyen_2006}. One of the problems of A* related approaches in RTS games is that moving units block/get in the way of/ each others and the path has to be recomputed often. Craig Reynolds \cite{Reynolds_1999} introduced algorithmic steering behaviors, which simulate the aggregate motion of swarms. His model describes how a system of multiple agents, each steering according to simple local rules, can produce the collective movement found in flocks of birds. Danielsiek et al. \cite{Danielsiek_2008} studied intelligent squad movements to do tactical moves to flank the opponent in a RTS game this way. An alternative to flocking is to use reciprocal velocity obstacles (RVO) \cite{Berg_2008}, since it is easier to control the behavior and they guarantee to prevent collisions. With this algorithm, we can move a large squad crossing another squad and avoid all possible collisions between units. {\color{blue}Do we put RVO? As I don't remember it being used in any game or game AI research.} In recent commercial RTS games like Starcraft 2 or Supreme Commander 2, flocking-like behaviors are inspired of continuum crowds (``flow field'') \cite{Treuille2006}, and a comprehensive review about (grid-based) pathfinding was recently done by Sturtevant \cite{sturtevant2012benchmarks}.

Another popular technique is potential fields (PF) and influence maps. Basically, PF work like a charged particle moving through a magnetic field. The parameters of PF charges can be learned through Q-Learning \cite{Liu_2008}. Another common issue of potential fields is the local optima problem that can stuck the unit away from the goal. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire, or staying at maximum shooting distance \cite{Hagelback08,Hagelback09}. Self-organizing-maps (SOM) learned with evolutionary algorithms provide some way to learn (realistic) parameters for RTS units behavior \cite{teamCompositionRTS}. Bayesian modeling has been applied to inverse fusion of the sensory inputs of the units \cite{SynnaeveMicroCig11}, which subsumes potential fields, allowing for integration of tactical goals directly in micro-management. There is also a cognitive approach to RTS AI \cite{SORTS}, using a human-like attention model (with units grouping and vision of a unique screen location) for micro-management.

There are some interesting uses of reinforcement learning (RL) \cite{Sutton}: Marthi et al. \cite{Marthi05} employ concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. Madeira et al. \cite{Madeira06} advocate the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. In real game setups, the actions spaces to explore is gigantic, it requires to use the structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. There are other way to learn parameters of an underlying model, for instance Ponsen and Spronck \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. The difficulty to work with multi-scale goals and plans is handled directly by case-based reasoning (CBR), which has been adapted for units behavior with continuous action models \cite{Molineaux08}, an integrated RL/CBR algorithm using continuous models, or with hybrid CBR/RL transfer learning \cite{CBR-RL}. Reactive planning \cite{WeberCig10}, a decompositional planning similar to hierarchical task networks \cite{HTNPlanning}, allows for plans to be changed at different granularity levels and so for multi-scale (hierarchical) goals integration of low-level control. Synnaeve and Bessi\`{e}re \cite{SynnaeveMicroCig11} achieve hierarchical goals (coming from tactical decisions) integration through the addition of another sensory input corresponding to the goal's objective.
